---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# autopca

<!-- badges: start -->
<!-- badges: end -->

Originally,this R package `autopca` was a script I used to draw PCA. PCA can analyze batch effects in experimental processing, and also analyze experimental processing factors. In my point, it is  very important in checking omics data. At present, this R package Features are very few But are under active development. The basic function is already there, so I placed on `github`, if necessary, you can download and use it yourself.

## Installation

The R packages only  can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("wangjiaxuan666/autopca")
```
## Example

Before we start the PCA analysis, we first need to tidy the input data. We can use the function `pca_data_tidy`to make the data clean for the next PCA analysis.It should be noted that the format of the input data is very important to determine the success of subsequent analysis.

Here, Emphasize the format of the input data. 
First, PCA analysis use the function `stat::prcomp`, Consequently the result given the every observed values'  Variance  in principal components. the observed values is the `prcomp` data' rownames.
For example, the data iris

```{r example}
head(iris)
```
the rownames is every iris ID number, in iris, we want to demonstrate the proportion of each iris flower in terms of petal length and width and so on.  That is we want, so for iris data, we just tidy a little for next analysis.
Just like

```{r}
library(autopca)
irisgroup <- iris[,5] 
iristidy <- iris[,-5]
pca(iristidy,sample_group = as.data.frame(irisgroup))
```
the group information is Grouping information is required, otherwise an error will be reported!

```{r}
#pca(iris)
```
As for why and how to use `autopca`, We need to start from the beginning.

## Illustration

the `autopca` designed for transcriptome data, The classic transcriptome data is rownames is gene id, the column names is every observed sample.  sometime it also add some annotation information in the tail. like this.

```{r}
test = matrix(rnorm(200), 20, 10)
test[1:10, seq(1, 10, 2)] = test[1:10, seq(1, 10, 2)] + 3
test[11:20, seq(2, 10, 2)] = test[11:20, seq(2, 10, 2)] + 2
test[15:20, seq(2, 10, 2)] = test[15:20, seq(2, 10, 2)] + 4
colnames(test) = paste("Test", 1:10, sep = "")
rownames(test) = paste("Gene", 1:20, sep = "")
annot <- c(rep("KEGG",20))
test <- data.frame(test,annot)
head(test)
```
Through the above steps, we obtained a classic transcriptome data frame. **NOW WE explain the sample variance in PC. So we need tidy the data by the function`pca_data_tidy`.**

```{r}
pca_data_tidy(as.data.frame(test)) -> test_tidy
```

NEXT, we can use the tidy data to analysis, just like this

```{r}
as.data.frame(c(rep("A",5),rep("B",5))) -> group
rownames(group) <- colnames(test[,-11])
colnames(group) <- "group"
head(group)
pca(test_tidy,sample_group = group)
```

Of course, as a fan of `tidyverse`, all function in `autopca` also  support `tibble` data input.
If there are any questions and suggestions in use, welcome questions and suggestions
